---
title: "Intro to R"
author: "Abhishek Bhatia"
date: "2023-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Iris Dataset

## Load in the Iris dataset

```{r}
library(datasets)
data(iris)

```

View the first ten rows of the Iris dataset using the `head()` function
```{r}
# use head()
```

Summarize all the variables using the `summary()` function
```{r}

```

## Tidyverse and the `dplyr` package

Read the collection of packages here: https://www.tidyverse.org/

```{r}
# Install the `tidyverse` package

# load `dplyr`
```

R is case-specific. Let's convert all the column names to lowercase to make this easier.
```{r}
names(iris) <- tolower(names(iris))
```


We will be using `dplyr` to manipulate data for now

### 1. `filter()`
```{r}
# filter() the data for species virginica.
virginica <- filter(iris, species == "virginica")

# View the first ten rows using head()
```

Now, try this on your own. Select only `virginica` species with `sepal.length > 5`
```{r}

```

### 2. Subsetting data using `select()`

Sometimes, data are "high dimensional". So, we only want a few number of columns, and we "subset" them. Try this yourself. 

```{r}
# Subset the iris dataset to only select sepal.length, sepal.width, and petal.length. Name the new dataframe "small_df"


```

R is great at understanding vectors. If I wanted to logically list out all numbers from 1-100, I can do the following:
```{r}
1:100
```
Similarly, it understands column names as a character vector. Try using this logic to select all columns from `sepal.length` to `petal length`.
```{r}

```


### 3. Changing variables with `mutate()`

`mutate()` allows you to create new variables, or change existing ones without changing much of the data around them. 
```{r}
# Create a new variable "greater.half"
iris <- mutate(iris, greater.half = sepal.width > 0.5 * sepal.length)

tail(iris)
```
Note the difference between `head()` and `tail()`

You can also use if/else statements to create "binary" or "dummy" variables as 0 and 1, instead of TRUE and FALSE, which is more useful for models. Try it below combining `mutate()` with `if_else()`

```{r}
# Mutate "greater.half" to now instead have 0 and 1 instead of TRUE and FALSE.

```

## Looking Ahead/Next time

Examining how variables are distributed
```{r}
hist(iris$sepal.width)
```
You can also look at how two variables are related
```{r}
plot(iris$sepal.width, iris$sepal.length)
```

